Dominance Regions for Imaginary $\bfg$-vectors

### 2021-08-28 dominance_regions.tex
Proof of dominant Dyck path recursion

@@

  \begin{lemma}
    Choose $a_1,a_2\in\ZZ_{\ge0}$.
    For $D^{st}:=D_1\sqcup D_2$ with $D_2=[1,a_2]$ and $D_1=[1,a_1]+a_2$, the element $x_D$ coincides with the corresponding standard monomial basis element.
  \end{lemma}

  Our main interest will be with the following Dyck paths.
  \begin{definition}
    \label{def:dominant Dyck path}
    For $a_1,a_2\in\ZZ_{\ge0}$ with $(-a_1+ba_2,-a_2)\in\cI$, define the \emph{dominant Dyck path} $D^{dom}=D^{dom}[a_1,a_2]$ to be the maximal Dyck path starting from $(0,0)$ and terminating at $(a_1,a_2)$ while staying below the lines $y=\frac{bc+\sqrt{bc(bc-4)}}{2b}x$ and $y-a_2=\frac{bc-\sqrt{bc(bc-4)}}{2b}(x-a_1)$.
  \end{definition}
  \begin{remark}
    \label{rem:intersection}
    The lines in the definition above meet at the point $\left( \frac{2ba_2-\big(bc-\sqrt{bc(bc-4)}\big)a_1}{2\sqrt{bc(bc-4)}} , \frac{\big(bc+\sqrt{bc(bc-4)}\big)a_2-2ca_1}{2\sqrt{bc(bc-4)}} \right)$.
  \end{remark}

@@

    For $t>0$, define the \emph{Dyck path suffix} $\overleftarrow{D}^t$ as the maximal Dyck path with $t$ horizontal edges starting from $(0,0)$ and terminating at $(t,h)$ while staying below the line $y-h=\frac{bc-\sqrt{bc(bc-4)}}{2b}(x-t)$, where $h=\left\lceil t \frac{bc-\sqrt{bc(bc-4)}}{2b}\right\rceil$.
    Call this $h$ the \emph{height} of $t$.
    %        y = (a2 - a1*(self._r1*self._r2-sqrt(self._r1*self._r2*(self._r1*self._r2-4)))/(2*self._r2))/(1-(self._r1*self._r2-sqrt(self._r1*self._r2*(self._r1*self._r2-4)))/(self._r1*self._r2+sqrt(self._r1*self._r2*(self._r1*self._r2-4))))
    %    corner = (y*(self._r1*self._r2-sqrt(self._r1*self._r2*(self._r1*self._r2-4)))/(2*self._r1),y)
  \end{definition}
  \begin{remark}
    There often exist multiple choices of $s$ with a given depth $d$ and multiple choices of $t$ with a given height $h$.
  \end{remark}

>>>>

  \begin{definition}
    Call a Dyck path $D$ 
    \begin{itemize}
      \item \emph{horizontally-adapted to $b$} if no vertical edge of $D$ is immediately preceded by more than $b$ horizontal edges;
      \item \emph{vertically-adapted to $c$} if no horizontal edge of $D$ is immediately followed by more than $c$ vertical edges.
    \end{itemize}
  \end{definition}

  \begin{definition}
    \label{def:Dyck path mutations}
    Given a Dyck path $D=D[a_1,a_2]$ which is horizontally-adapted to $b$, define a Dyck path $\mu_{V,b}(D)$ inside the lattice rectangle $[0,a_2]\times[0,ba_2-a_1]$ by replacing each vertical edge, together with the $\ell$ horizontal edges which immediately precede it, by a horizontal edge followed by $b-\ell$ vertical edges.

    Given a Dyck path $D=D[a_1,a_2]$ which is vertically-adapted to $c$, define a Dyck path $\mu_{H,c}(D)$ inside the lattice rectangle $[0,ca_1-a_2]\times[0,a_1]$ by replacing each horizontal edge, together with the $\ell$ vertical edges which immediately follow it, by $c-\ell$ horizontal edges followed by a vertical edge.
  \end{definition}
  \begin{remark}
    Observe that $\mu_{H,a}$ and $\mu_{V,a}$ are inverse for any $a\ge1$.
  \end{remark}

  \begin{lemma}\mbox{}
    \begin{enumerate}
      \item If $a_1 < b a_2$, then $\mu_{V,b}\big(D^{max}[a_1,a_2]\big)=D^{max}[a_2,ba_2-a_1]$.
      \item If $a_2 < c a_1$, then $\mu_{H,c}\big(D^{max}[a_1,a_2]\big)=D^{max}[ca_1-a_2,a_1]$.
    \end{enumerate}
  \end{lemma}
  \begin{proof}
    The proof is identical to the proof of \cite[Lemma 2.4]{rupel} with appropriate notation changes.
  \end{proof}

@@

  \begin{lemma}
    Choose $a_1,a_2\in\ZZ_{\ge0}$.
    For $D^{st}:=D_1\sqcup D_2$ with $D_2=[1,a_2]$ and $D_1=[1,a_1]+a_2$, the element $x_{D^{st}}$ coincides with the corresponding standard monomial basis element.
  \end{lemma}

  Our main interest will be with the following Dyck paths.
  \begin{definition}
    \label{def:dominant Dyck path}
    For $a_1,a_2\in\ZZ_{\ge0}$ with $(-a_1+ba_2,-a_2)\in\cI$, define the \emph{dominant Dyck path} $D^{dom}=D^{dom}[a_1,a_2]$ to be the maximal Dyck path starting from $(0,0)$ and terminating at $(a_1,a_2)$ while staying below the lines $y=\frac{bc+\sqrt{bc(bc-4)}}{2b}x$ and $y-a_2=\frac{bc-\sqrt{bc(bc-4)}}{2b}(x-a_1)$.

    For $a_1,a_2\in\ZZ_{\ge0}$ with $(-a_2,-a_1+ca_2)\in\cI'$, define the \emph{dual dominant Dyck path} $D^{dom^\vee}=D^{dom^\vee}[a_1,a_2]$ to be the maximal Dyck path starting from $(0,0)$ and terminating at $(a_1,a_2)$ while staying below the lines $y=\frac{bc+\sqrt{bc(bc-4)}}{2c}x$ and $y-a_2=\frac{bc-\sqrt{bc(bc-4)}}{2c}(x-a_1)$.
  \end{definition}
  \begin{remark}
    \label{rem:intersection}
    The lines in the definition above meet at the point $\left( \frac{2ba_2-\big(bc-\sqrt{bc(bc-4)}\big)a_1}{2\sqrt{bc(bc-4)}} , \frac{\big(bc+\sqrt{bc(bc-4)}\big)a_2-2ca_1}{2\sqrt{bc(bc-4)}} \right)$.
  \end{remark}

@@

    For $t>0$, define the \emph{Dyck path suffix} $\overleftarrow{D}^t$ as the maximal Dyck path with $t$ horizontal edges starting from $(0,0)$ and terminating at $(t,h)$ while staying below the line $y-h=\frac{bc-\sqrt{bc(bc-4)}}{2b}(x-t)$, where $h=\left\lceil t \frac{bc-\sqrt{bc(bc-4)}}{2b}\right\rceil$.
    Call this $h$ the \emph{height} of $t$.
  \end{definition}
  \begin{remark}
    There often exist multiple choices of $s$ with a given depth $d$ and multiple choices of $t$ with a given height $h$.
  \end{remark}

  \begin{lemma}
    For any $s,t>0$, the Dyck path prefix $\overrightarrow{D}^s$ and Dyck path suffix $\overleftarrow{D}^t$ are both horizontally-adapted to $b$ and vertically-adapted to $c$.
  \end{lemma}

  \begin{theorem}
    For $a_1,a_2\in\ZZ_{\ge0}$ with $(-a_1+ba_2,-a_2)\in\cI$, we have $\mu_{V,b}\big(D^{dom}[a_1,a_2]\big)=D^{dom^\vee}[a_2,ba_2-a_1]$ and $\mu_{H,c}\big(D^{dom}[a_1,a_2]\big)=D^{dom^\vee}[ca_1-a_2,a_1]$.
  \end{theorem}
  \begin{proof}
    We prove the result for $\mu_{H,c}$ as the result for $\mu_{V,b}$ immediately follows from $\mu_{H,b}$ being its inverse.
    Write $D'$ for the Dyck path $\mu_{H,c}\big(D^{dom}[a_1,a_2]\big)$ and observe that by construction $D'$ has $ca_1-a_2$ horizontal edges and $a_1$ vertical edges.
    Write $v'_1,\ldots,v'_{a_1}$ for the vertical edges of $D'$ with $v'_r$ immediately preceded by $\ell_r$ horizontal edges.
    We have to show (i) that every lattice point on $D'$ stays below the lines 
    \[y=\frac{bc+\sqrt{bc(bc-4)}}{2c}x \qquad \text{and} \qquad y-a_1=\frac{bc-\sqrt{bc(bc-4)}}{2c}\big(x-(ca_1-a_2)\big)\]
    while (ii) any lattice point above $D'$ is also above one of these lines.

    If there exists $t$ so that $v'_t$ passes above the line $y=\frac{bc+\sqrt{bc(bc-4)}}{2c}x$, then
    \[ \frac{t}{\sum\limits_{r=1}^t \ell_r} > \frac{bc+\sqrt{bc(bc-4)}}{2c} \]
    or, rearranging, we have
    \[ \frac{\sum\limits_{r=1}^t \ell_r}{t} < \frac{2c}{bc+\sqrt{bc(bc-4)}} = \frac{bc-\sqrt{bc(bc-4)}}{2b} \]
    or
    \[ \frac{bc+\sqrt{bc(bc-4)}}{2b} < \frac{ct-\sum\limits_{r=1}^t \ell_r}{t} = \frac{\sum\limits_{r=1}^t (c-\ell_r)}{t}. \]
    But this states that the subpath of $D^{dom}[a_1,a_2]$ consisting of the first $t$ horizontal edges and all their trailing vertical edges passes above the line $y=\frac{bc+\sqrt{bc(bc-4)}}{2b}x$, contrary to the definition.

    If there exists $t$ so that $v'_t$ passes above the line $y-a_1=\frac{bc-\sqrt{bc(bc-4)}}{2c}\big(x-(ca_1-a_2)\big)$, then
    \[ \frac{bc-\sqrt{bc(bc-4)}}{2c} > \frac{a_1-t}{ca_1-a_2-\sum\limits_{r=1}^t \ell_r} \]
    or, rearranging, we have
    \[ \frac{ca_1-a_2-\sum\limits_{r=1}^t \ell_r}{a_1-t} > \frac{2c}{bc-\sqrt{bc(bc-4)}} = \frac{bc+\sqrt{bc(bc-4)}}{2b} \]
    or
    \[ \frac{bc-\sqrt{bc(bc-4)}}{2b} > \frac{ c(a_1-t) - \big(ca_1-a_2-\sum\limits_{r=1}^t \ell_r\big)}{a_1-t} = \frac{a_2-\sum\limits_{r=1}^t (c-\ell_r)}{a_1-t}. \]
    But this states that the subpath of $D^{dom}[a_1,a_2]$ consisting of the first $t$ horizontal edges and all their trailing vertical edges passes above the line $y-a_2=\frac{bc-\sqrt{bc(bc-4)}}{2b}(x-a_1)$, contrary to the definition.

    This proves (i), analogous calculations establish (ii). 
  \end{proof}
